On The Impact of Atoms of Confusion in JavaScript Code

### 2022-03-31 3_Method.tex
Fixed some terminology issues about obfuscated and clean versions where we were employing different terms.

By answering these research questions, we can either generalize or refute the findings about atoms of confusion already discussed in the literature (goal of research questions (RQ1) and (RQ2)). In addition, answer to the third question allows us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice\castor{This should probably be removed.}. Answering these research questions also lays the foundations for the implementation of tools that can automatically transform code into cleaner versions.
%---though we postpone these results to future research work.
We conduct a mixed-methods study to answer these questions. It includes two independently designed and conducted experiments\castor{I would rather use the term ``experiment''.}: a replication of the first study on atoms of confusion, but targeting JavaScript programs, and a new study using a latin square counterbalancing experimental design to control for subject experience. 
These two experiments explore the impact of atom candidates on understanding JavaScript code (goal of research question RQ1). They investigate whether or not programs that contain atom candidates produce a higher misunderstanding rate when programmers try to predict their outcome.\castor{I've removed the part about time because, in order for the first study to be a replication, we need to compute time differently, accounting for incorrect responses as well.}% and (b) require more time for programmers to predict their output. 
We leverage two independently-developed studies with different designs to more rigorously verify the impact of atoms of confusion in code comprehension activities. 
We also conduct a set of interviews with developers to contrast the quantitative results of the experiments with their preferences and opinions (thus addressing RQ2).

@@

The study is conducted by means of a web application. As part of this effort, we carried out an informal pilot whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the survey would take on average. Based on the feedback of the participants of the pilot, we present only one obfuscated and one clean version of each atom candidate, to reduce to total time of the study and potentially increase participation. Initially, our plan was to present two of each.

@@

This experiment complements the replication study by using a different experimental design and controlling for one additional variable, the experience of the subjects. For this study we employ a subset of the atom candidates used in the replication study. More specifically, we
%Since JavaScript and \clang have some constructs in common (Section~\ref{back}), we first 
select a set of 9+1 atom candidates: nine previously-validated atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} that also exist in JavaScript programs plus one atom candidate that is specific to the JavaScript language (Automatic Semicolon Insertion)\castor{I think the name of this atom is inconsistent between the studies}. 
Our supplementary material discusses and shows examples of the atoms we consider in our research.\castor{We should probably do this here or just refer to the table that I intend to insert for the replication study.}

@@

The design of our second study blocks two variables (subject experience and the programs) and
considers two treatments: the presence or absence of atom candidates within the programs. 
To achieve such a design goal, controlling the effect of experience and individual programs, we resort to the \textit{Latin Square Design} \cite{Hunter-Experimenters}. Using this design we create a 2 x 2 matrix in which each row represents a subject and each column indicates the set of programs. The design of each square (a replica) is such that no treatment is repeated in the same row or column. For example, considering that we have a set of 10
code samples $s_0, s_1, ..., s_{10}$, if a given subject (P1) is asked to predict the output of the code samples $s_0, s_1, ..., s_5$ that contain atom candidates, then, when answering questions about non-confusing programs, they will only be presented with non-confusing versions of the code samples $s_6, s_7,..., s_{10}$. Furthermore, a given subject (P2), who constitutes the second row of our example square, will be asked questions about the non-confusing versions for $s_1, s_2, ..., s_5$, and will answer questions about confusing programs for $s_6, s_7,..., s_{10}$. By doing that, we guarantee that all versions of the programs are contained within each square, and that each configuration occurs only once within a square. Figure \ref{fig:latinsquare} offers a visual representation of the concept.

@@

For each of the 10 selected atom candidates, we write one short program containing the atom. As in the replication study, we call them the confusing versions of the programs. We also write corresponding short programs without the atom candidates and call them the clean versions of the programs. Overall, the experiment employed 20 programs, 10 confusing and 10 clean. In order to reduce the cognitive effort, each subject was asked to predict the output of 10 programs, five confusing and five clean. The order in which they are presented is randomized. By doing this, we seek to minimize the chances of subjects being aware that the current listing they are analyzing contains (or not) atoms of confusion. That is, each subject should indicate what would be the outcomes of the programs, some of them having atoms of confusion (while other programs do not). Since participants are not exposed to obfuscated and clean versions corresponding to the same atom candidate, learning effect is not possible. We measure answer correctness and the total time each participant requires to participate in the experiment.

@@

We posted the survey on a JavaScript Reddit channel.\footnote{https://www.reddit.com/r/javascript/} We explained our research purposes and asked developers of any level of expertise to take the survey. Within twelve hours, we collected more than 150 answers, populating more than 70 replicas of the Latin Squares. We collected significant data on time taken and discrepancies in answer correctness between confusing and non-confusing versions of the programs. Some inconsistencies arose while building the squares, for instance, when a user quit in the middle of the survey. We discarded from our study all squares that contained incomplete rows (a common approach for data imputation).
%Since we had a large enough number of samples, the squares we had to discard did not impact our analysis.

@@

In the third part of the interview, participants are shown pairs of programs that were used in the first study, where each pair contains a clean and a confusing version of the same atom candidate. The participants 
%were told both sides of each pair had the same output, and 
are asked to evaluate which version of the code is easier to understand. To avoid introducing bias in the answers, the interviewers do not explain that one of the versions in each pair contains the atom under investigation. Subjects are just presented the pairs of programs and allowed to take the necessary time to decide on the most readable one.

>>>>

By answering these research questions, we can either generalize or refute the findings about atoms of confusion already discussed in the literature (goal of research questions (RQ1) and (RQ2)). In addition, answering the third question allows us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice\castor{This should probably be removed.}. Answering these research questions also lays the foundations for the implementation of tools that can automatically transform code into cleaner versions.
%---though we postpone these results to future research work.
We conduct a mixed-methods study to answer these questions. It includes two independently designed and conducted experiments\castor{I would rather use the term ``experiment''.}: a replication of the first study on atoms of confusion, but targeting JavaScript programs, and a new study using a latin square counterbalancing experimental design to control for subject experience. 
These two experiments explore the impact of atom candidates on understanding JavaScript code (goal of research question RQ1). They investigate whether or not programs that contain atom candidates produce a higher misunderstanding rate when programmers try to predict their outcome.\castor{I've removed the part about time because, in order for the first study to be a replication, we need to compute time differently, accounting for incorrect responses as well.}% and (b) require more time for programmers to predict their output. 
We leverage two independently-developed studies with different designs to more rigorously verify the impact of atoms of confusion in code comprehension activities. 
We also conduct a set of interviews with developers to contrast the quantitative results of the experiments with their preferences and opinions (thus addressing RQ2).

@@

The study is conducted by means of a web application. As part of this effort, we carried out an informal pilot whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the survey would take on average. Based on the feedback of the participants of the pilot, we present only one obfuscated and one clean version of each atom candidate, to reduce to total time of the study and potentially increase participation.% Initially, our plan was to present two of each.

@@

This experiment uses a different experimental design that makes it possible to control for the experience of the subjects. For this study we employ a subset of the atom candidates used in the replication study. More specifically, we
%Since JavaScript and \clang have some constructs in common (Section~\ref{back}), we first 
select a set of 9+1 atom candidates: nine previously-validated atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} that also exist in JavaScript programs plus one atom candidate that is specific to the JavaScript language (Automatic Semicolon Insertion)\castor{I think the name of this atom is inconsistent between the studies}. 
Our supplementary material discusses and shows examples of the atoms we consider in our research.\castor{We should probably do this here or just refer to the table that I intend to insert for the replication study.}

@@

The design of our second study blocks two variables (subject experience and the programs) and
considers two treatments: the presence or absence of atom candidates within the programs. 
To achieve such a design goal, controlling the effect of experience and individual programs, we resort to the \textit{Latin Square Design} \cite{Hunter-Experimenters}. Using this design we create a 2 x 2 matrix in which each row represents a subject and each column indicates the set of programs. The design of each square (a replica) is such that no treatment is repeated in the same row or column. For example, considering that we have a set of 10
code samples $s_0, s_1, ..., s_{10}$, if a given subject (P1) is asked to predict the output of the code samples $s_0, s_1, ..., s_5$ that contain atom candidates, then, when answering questions about clean programs, they will only be presented with clean versions of the code samples $s_6, s_7,..., s_{10}$. Furthermore, a given subject (P2), who constitutes the second row of our example square, will be asked questions about the clean versions for $s_1, s_2, ..., s_5$, and will answer questions about obfuscated programs for $s_6, s_7,..., s_{10}$. By doing that, we guarantee that all versions of the programs are contained within each square, and that each configuration occurs only once within a square. Figure \ref{fig:latinsquare} offers a visual representation of the concept.

@@

For each of the 10 selected atom candidates, we write one short program containing the atom. As in the replication study, we call them the obfuscated versions of the programs. We also write corresponding short programs without the atom candidates and call them the clean versions of the programs. Overall, the experiment employed 20 programs, 10 obfuscated and 10 clean. In order to reduce the cognitive effort, each subject was asked to predict the output of 10 programs, five obfuscated and five clean. The order in which they are presented is randomized. By doing this, we seek to minimize the chances of subjects being aware that the current listing they are analyzing contains (or not) atoms of confusion. That is, each subject should indicate what would be the outcomes of the programs, some of them having atoms of confusion (while other programs do not). Since participants are not exposed to obfuscated and clean versions corresponding to the same atom candidate, learning effect is not possible. We measure answer correctness and the total time each participant requires to participate in the experiment.

@@

We posted the survey on a JavaScript Reddit channel.\footnote{https://www.reddit.com/r/javascript/} We explained our research purposes and asked developers of any level of expertise to take the survey. Within twelve hours, we collected more than 150 answers, populating more than 70 replicas of the Latin Squares. We collected significant data on time taken and discrepancies in answer correctness between obfuscated and clean versions of the programs. Some inconsistencies arose while building the squares, for instance, when a user quit in the middle of the survey. We discarded from our study all squares that contained incomplete rows (a common approach for data imputation).
%Since we had a large enough number of samples, the squares we had to discard did not impact our analysis.

@@

In the third part of the interview, participants are shown pairs of programs that were used in the first study, where each pair contains a clean and an obfuscated version of the same atom candidate. The participants 
%were told both sides of each pair had the same output, and 
are asked to evaluate which version of the code is easier to understand. To avoid introducing bias in the answers, the interviewers do not explain that one of the versions in each pair contains the atom under investigation. Subjects are just presented the pairs of programs and allowed to take the necessary time to decide on the most readable one.

### 2022-03-31 discussion.tex
remocao de texto nao essencial para ajustar as limitacoes de espaco
Fixed some terminology issues about obfuscated and clean versions where we were employing different terms.

Our work has several implications.
First, it adds external validity to the
work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17},
which investigates
the impact of atom candidates on
understanding \clang,\cpplang code. That is,
similarly to their work, the atom candidates
Comma Operator, Post/Pre Increment, Omitted Curly Braces,
Assignment as Value, Implicit Predicate, and Logic as
Control Flow seem to make 
JavaScript code harder to understand. For five of them, the difference is statistically significant, with a large effect size for four atoms. Our results also 
refute the hypothesis that Arithmetic as Logic is an atom of confusion (i.e., a source of misunderstanding).
In comparison to the original
work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, 
our study
led to some differences in the effect size
of the atom candidates.
Altogether, we answer our first research question
\emph{\rqa}:

@@

The results of the interview study complement the understanding of atoms of confusion because the participants make clear the existence of a trade-off between code comprehension and other quality attributes. For instance, most of the participants prefer the version of the code with the Conditional Operator, even though they agree that its use might contribute to the misunderstanding of JavaScript code, particularly when novices are maintaining the codebase. The participants of the interview study also
mentioned other possible sources of misunderstanding in JavaScript,
including the use of prototype-based inheritance and nested call-backs (as discussed in Section~\ref{sec:interview-results}). Other JavaScript atom candidates include
Object Destructuring, Array Spread, Object Spread, and Type Conversion.
In summary, the results of the second study (interviews) allow
us to answer the second (\rqb) research question.

@@

The results of the third study (mining open source
JavaScript repositories) provides evidence that,
although atoms of confusion compromise program
comprehension, they frequently appear in open
source JavaScript projects. In particular,
seven, out of 10 atoms considered
  in our study, appear in more than 50\% of
the 72 projects we analyzed. Furthermore, at least two of them are used intensively, more than once for every 200 lines of code. In summary, the third study
allows us to answer the fourth research
question (\emph{\rqd}).

\begin{mh}
  {\bf Answer to RQ3:} The MSR study reveals that
  the several atom candidates explored in our research
  appear frequently in practice, and cleaning up the use of 
  Post-Increment/Decrement and the Automatic Semicolon Insertion
  might improve the readabiliy of JavaScript code substantially. 
(\emph{\rqc}) 
\end{mh}

>>>>

Our work has several implications.
First, it adds external validity to the
work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17},
which investigates
the impact of atom candidates on
understanding \clang,\cpplang code. That is,
similarly to their work, the atom candidates
Comma Operator, Post/Pre Increment, Omitted Curly Braces,
Assignment as Value, Implicit Predicate, and Logic as
Control Flow seem to make 
JavaScript code harder to understand. For five of them, the difference is statistically significant, with a large effect size for four atoms. Our results also 
refute the hypothesis that Arithmetic as Logic is an atom of confusion (i.e., a source of misunderstanding).
In comparison to the original
work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, 
our study
led to some differences in the effect size
of the atom candidates.
Altogether, we answer our first research question.

@@

The results of the interview study complement the understanding of atoms of confusion because the participants make clear the existence of a trade-off between code comprehension and other quality attributes. For instance, most of the participants prefer the version of the code with the Conditional Operator, even though they agree that its use might contribute to the misunderstanding of JavaScript code, particularly when novices are maintaining the codebase. The participants of the interview study also
mentioned other possible sources of misunderstanding in JavaScript,
including the use of prototype-based inheritance and nested call-backs (as discussed in Section~\ref{sec:interview-results}). Other JavaScript atom candidates include
Object Destructuring, Array Spread, Object Spread, and Type Conversion.
In summary, the results of the second study (interviews) allow
us to answer the second research question.

@@

The results of the third study (mining open source
JavaScript repositories) provides evidence that,
although atoms of confusion compromise program
comprehension, they frequently appear in open
source JavaScript projects. In particular,
seven, out of 10 atoms considered
  in our study, appear in more than 50\% of
the 72 projects we analyzed. Furthermore, at least two of them are used intensively, more than once for every 200 lines of code. In summary, the third study
allows us to answer the third research
question.

\begin{mh}
  {\bf Answer to RQ3:} The MSR study reveals that
  the several atom candidates explored in our research
  appear frequently in practice, and cleaning up the use of 
  Post-Increment/Decrement and the Automatic Semicolon Insertion
  might improve the readabiliy of JavaScript code substantially. 
\end{mh}

### 2022-03-31 main.tex
Fixed some terminology issues about obfuscated and clean versions where we were employing different terms.

\begin{abstract}
  Evolving legacy code is a challenging task, particularly when the code has been poorly written or uses confuse idioms and language constructs, which might increase maintenance efforts and impose a significant cognitive load on developers. For this reason, researchers have investigated possible sources of confusion in codebases, including the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in statically typed languages such as \clang, \cpplang, and Java. In this work, we investigate whether atoms of confusion identified in statically typed languages also confuse developers of a dynamically typed language. We use JavaScript as a representative example of dynamic programming languages and collect evidence from a mixed-method research effort: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) study. Our survey and interviews confirm that atoms of confusion lead to code that is hard to understand in JavaScript. Considering ten atom candidates, developers correctly predict the outcome of at least 15\% more cases in code snippets where atom candidates are not present, compared to alternative versions that include the atom candidates. For five of these atoms, the difference is statistically significant (accounting for p-value correction). Effect size is large for four atoms and medium for one of them. In addition, our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems. Four atom candidates appear in 90\% of the analyzed projects, and two of them occur more than once for every 100 lines of code in the dataset. Altogether, our findings might help practitioners: (1) better understand the implications of atoms of confusion on understanding JavaScript code, (2) avoid writing code that is unnecessarily difficult to maintain, and (3) design program transformation tools that remove potential sources of misunderstanding in JavaScript code.
\end{abstract}

>>>>

\begin{abstract}
  Evolving legacy code is a challenging task, particularly when the code has been poorly written or uses confuse idioms and language constructs, which might increase maintenance efforts and impose a significant cognitive load on developers. For this reason, researchers have investigated possible sources of confusion in codebases, including the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in statically typed languages such as \clang, \cpplang, and Java. In this work, we investigate whether atoms of confusion identified in statically typed languages also confuse developers of a dynamically typed language. We use JavaScript as a representative example of dynamic programming languages and collect evidence from a mixed-method research effort: a survey, a set of interviews with practitioners, and an activity of mining open source JavaScript repositories (MSR). Our survey and interviews confirm that atoms of confusion lead to code that is hard to understand in JavaScript. Considering ten atom candidates, developers correctly predict the outcome of at least 15\% more cases in code snippets where atom candidates are not present, compared to alternative versions that include the atom candidates. For five of these atoms, the difference is statistically significant (accounting for p-value correction). Effect size is large for four atoms and medium for one of them. In addition, our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems. Four atom candidates appear in 90\% of the analyzed projects, and two of them occur more than once for every 100 lines of code in the dataset. Altogether, our findings might help practitioners: (1) better understand the implications of atoms of confusion on understanding JavaScript code, (2) avoid writing code that is unnecessarily difficult to maintain, and (3) design program transformation tools that remove potential sources of misunderstanding in JavaScript code.
\end{abstract}

### 2022-03-31 method_replication_study.tex
Fixed some terminology issues about obfuscated and clean versions where we were employing different terms.

The study is conducted by means of a web application. We conducted an initial pilot to find bugs with the application and to get general impressions. Based on the feedback of the participants of the pilot, we decided to present only one obfuscated and one clean version of each atom candidate, to reduce to total time of the study and potentially increase participation. Initially, our plan was to present two of each. 
To reduce the likelihood of a subject simply running the program, the web application verifies whether the subject attempts to change tabs or windows and presents the programs as images, instead of text.

We organize the survey in three sections. In the first one we present instructions and also a check button whose checking meant users agreed that all collected data would be used anonymously and solely for research purposes. The instructions explain how the survey works and asks them to dedicate their attention to it.  We stressed to participants the importance of not using any aids during the survey, such as online or console interpreters. For each question page, we kept track of whether or not the subjects switched windows.

The last section of the survey presented a sequence of ten questions, each containing a code snippet. For each question, there was a text box where the answer should be written. There was also an ``I do not know'' button, which, when clicked, led the subject to the next question. In our setting, ``I do not know'' was treated as a wrong answer. The code snippets were presented as images copied from a text editor, so as to demotivate respondents from resorting to external resources by copying and pasting the code into an interpreter. Upon submitting their answer for a particular question, the subject was automatically led to a similar page, containing another snippet.

We decided not to provide feedback about the time subjects took to answer each question. We also did not tell them whether their answers were correct or not. Our main concern was to avoid introducing bias for future respondents. Since we posted the survey on a social media platform, possible threats could have arisen if we gave respondents instant feedback.

>>>>

The study is conducted by means of a web application. We carried out an initial pilot to find bugs with the application and to get general impressions. Based on the feedback of the participants of the pilot, we decided to present only one obfuscated and one clean version of each atom candidate, to reduce to total time of the study and potentially increase participation. Initially, our plan was to present two of each.

We organize the study in three sections. In the first one we present instructions and also a check button whose checking meant users agreed that all collected data would be used anonymously and solely for research purposes. The instructions explain how the survey works and asks them to dedicate their attention to it.  We stressed to participants the importance of not using any aids during the survey, such as online or console interpreters.

The second section presents the programs, one at a time. For each one, there is a text box where the answer should be written. There is also an ``I do not know'' button, which, when clicked, leads the subject to the next question. In our setting, ``I do not know'' was treated as a wrong answer. To reduce the likelihood of a subject leveraging attempting to execute the programs, the web application verifies whether the subject attempts to change tabs or windows and presents the programs as images, instead of text. Upon submitting an answer for a particular program, the subject is automatically led to a similar page, containing the next one.

We do not provide feedback about the time subjects take to answer each question. We also do not tell them whether their answers are correct or not. This aims at avoiding introducing bias for future respondents. Since we posted the survey on social media platforms, possible threats could have arisen if we gave respondents instant feedback.
